GSE158420: RNA expression downsampling

### 2024-07-17 analysis/GSE158420_rna.Rmd
mds looks okay
show data
add data
data
thincounts

## Load annotations and methylation data

@@

@@

```{r,deseq2}

RhpcBLASctl::blas_set_num_threads(4) #set automatic cores to 1

@@

@@

@@

This will run the LAM analysis at group sizes of 2, 3, 6, 10, 16, 22 and 30,
50 times each, with pseudorandom seeds from 100 to 5000 jumping up by 100.

@@

@@

@@

```
## Session information

>>>>

## Load pathway data

@@

MDS plot

```{r,mds}

mds <- cmdscale(dist(t(scale(x))))

cols <- gsub("1","pink",gsub("0","lightblue",tumor))

plot(mds, xlab="Coordinate 1", ylab="Coordinate 2", pch=16,col=cols,cex=2)

text(mds, labels=colnames(x),cex=0.7)

@@

```{r,deseq2}
RhpcBLASctl::blas_set_num_threads(4) #set automatic cores to 1

@@

str(fsets)

@@

# Full counts

@@

This will run the FGSEA analysis at group sizes of 2, 3, 6, 10, 16, 22 and 30,
50 times each, with pseudorandom seeds from 100 to 5000 jumping up by 100.

@@

dots <- lapply(ssres,function(x) { x[,"TP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fgsea",ylim=c(0,nrow(fsig))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=nrow(fsig))

dots <- lapply(ssres,function(x) { x[,"FP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

@@

dots <- lapply(ssres,function(x) { x[,"TP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fora",ylim=c(0,length(osets))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=length(osets))

dots <- lapply(ssres,function(x) { x[,"FP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

@@

# Thin Counts

```{x,thincounts}

x2 <- thinCounts(x, prob=NULL, target.size=10000000)
colSums(x2)
x <- x2

This will run the FGSEA analysis at group sizes of 2, 3, 6, 10, 16, 22 and 30,
50 times each, with pseudorandom seeds from 100 to 5000 jumping up by 100.

```{r,fgsea_run2,fig.height=10}

pdf("sens_fgsea_thin.pdf",height=7,width=5)
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fgsea",ylim=c(0,nrow(fsig))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=nrow(fsig))

dots <- lapply(ssres,function(x) { x[,"FP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

```{r,fora_run2,fig.height=10}

pdf("sens_fora_thin.pdf",height=7,width=5)
par(mfrow=c(3,1))

dots <- lapply(ssres,function(x) { x[,"TP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. consistent",xlab="sample size",cex=0,main="fora",ylim=c(0,length(osets))) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)
abline(h=length(osets))

dots <- lapply(ssres,function(x) { x[,"FP"] })
boxplot(dots,names=SAMPLESIZES,col="white",ylab="no. inconsistent",xlab="sample size",cex=0) ; grid()
beeswarm(dots,add=TRUE,pch=1,cex=0.6)

dots <- lapply(ssres,function(x) { x[,"TP"] })
TPdots <- do.call(rbind,dots)
dots <- lapply(ssres,function(x) { x[,"FP"] })
FPdots <- do.call(rbind,dots)
ratio <- t(FPdots / (TPdots + FPdots))
ratio <- lapply(1:ncol(ratio),function(i) {ratio[,i]})
boxplot(ratio,names=SAMPLESIZES,col="white",ylab="proportion inconsistent",xlab="sample size",cex=0,ylim=c(0,0.5)) ; grid()
beeswarm(ratio,add=TRUE,pch=1,cex=0.6)

### PDF
pdf("sens_thin.pdf",height=4,width=8)
par(mfrow=c(1,3))

## Session information
